Argennon: A Scalable Smart Contract Platform

### 2022-03-28 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The AscEE has two type of execution sessions: optimistic and monitored. Resource usage of an optimistic session
is always measured per session and default pre-defined resource caps are used. An optimistic session must always
complete normally. On the other hand in a monitored session some resources are measured per application and some
caps are determined dynamically by the request. A monitored session is allowed to either complete normally or fail.

Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise measurement of
used resources. Note that for implementing optional decoupling principle we need to determine the exact application
which fails in the call chain. For example assume that an optimistic session containing an application call violates a
1ms execution time cap. If the call happens exactly after 1ms a small fluctuation in measuring the execution time can
change the point of failure between the called and the main application. On the other hand this is not an issue for
monitored session because resource measurements and caps are defined per application.

The following resources are defined and monitored for AscEE execution sessions:

execution time:
The amount of cpu time that is required for executing a session or an application. The execution time of a session
is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
of cpu time needed for executing a predefined standard application which is used for benchmarking a host's performance.

local memory: any memory usage of an application that is not part of the heap chunks and is not part of another resource
will be considered as local memory usage. This resource is measured in bytes.

heap access list: every session can only access heap locations that are declared in its access list. In addition,
resizing
heap chunks can only be done in the range of the pre-declared lower bound and upper bound. those predeclared
locations of heap chunks that are

applications list: a session can only make requests to applications that are declared in its application list.

call depth: the number of nested application calls.

number of deferred calls
number of session locks
number of virtual signatures

>>>>

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session some resources are measured per application and some
caps are determined dynamically by the request. An optimistic session must always
complete normally but a monitored session is allowed to either complete normally or abruptly.

Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise resource
measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
which has failed in the call chain. For example assume that an optimistic session containing an application call
violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
milliseconds of execution, a small fluctuation in the execution time measurement can
change the point of failure between the called and the main application. This will introduce nondeterministic
behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
because resource measurements and caps are defined per application, and we make sure that resource caps are not too
small for any application call.

Different computational resources are measured and monitored during an AscEE session:
\begin{itemize}
    \item \textbf{execution time}:
    the amount of cpu time that is required for executing a session or an application. The execution time of a session
    is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
    of cpu time needed for executing a predefined standard application which is used for benchmarking a host's
    performance.

    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
    value defines a bound on the total cpu time of the session and no per application measurement is done.

    Monitored sessions perform per application cpu-time measurement, and every application call during a monitored
    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
    can be used for executing that particular application call. It should be noted that the timer is paused when an
    application makes a call to another application, until the control returns to the caller. An application call
    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
    considered a failed call.

    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
    application it has to \emph{forward} a portion of its external clocks to the called application. This amount
    will determine the value of maxClocks for the called application.

    The amount of external clocks for every application
    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result the total number of clocks of a monitored
    session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
    \texttt{maxClocks} for
    the root application call is determined by the transaction.

    \item \textbf{local memory}:
    any memory usage of an application that is not part of the heap chunks and is not part of another resource
    will be considered as local memory usage. This resource is measured in bytes.
    \item \textbf{heap access list}:
    every session can only access heap locations that are declared in its access list. In addition,
    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
    \item \textbf{applications list}:
    a session may only make requests to applications that are declared in its application list.
    \item \textbf{call depth}:
    the number of nested application calls can not be more than a threshold. This threshold is determined by the
    Argennon protocol.
    \item active differed calls:
    \item virtual signatures:
    \item number of locks

\end{itemize}
